Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-10-01 files/Chapter6/Validity_2D.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of the Timoshenko beam model}
In the article \cite{LVV09}, the authors show that a cantilever Timoshenko beam model can be a feasible model to replace a more realistic two-dimensional model for use in application. They also show that the dimensions of the beam will determine how well the one-dimensional Timoshenko beam compares to the two-dimensional beam. In this section, the results of the article is discussed, replicated and extended with more testcases.

Where \cite{LVV09} differs from most other research into the validity of the Timoshenko beam theory, is that the boundary conditions are also taken into consideration. The article serves as an intermediate step to comparing a Timoshenko beam to a three-dimensional beam.

The authors use a cantilever beam with rectangular cross-section.

\subsection{The models}
Recall the cantilever Timoshenko beam model from section \ref{ssec:1D_Model:ModelProblems}, referred to as Problem T-2 and the cantilever two-dimensional beam model from section \ref{ssec:2D_Model:ModelProblem}, referred to as Problem 2D-1. In \cite{LVV09} it is assumed that the beams have a rectangular cross-section.\\

Figure 6.1 shows the two beams side-by-side.

\begin{figure}[h!]\label{fig:compare:1D+2D}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side by side comparison of the beams.}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
					\draw[line width = 0.4mm] (0,0) -- (6,0);

					\node at (2.7,-0.25) {$\ell = 1$};

					\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
					\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
					\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
					\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
					\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
					\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
					\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
					\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
					\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

					\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
					\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
					\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
					\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
					\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
					\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
					\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
					\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

					%\node at (3.2,-0.2) {$\ell = 1$};
					\end{tikzpicture}
				\end{center}
				\subcaption{Timoshenko Cantilever Beam}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
						\draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
						\draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
						\draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);
						%\draw[line width = 0.4mm] (0,-0.2) -- (0,0.2);

						%\draw[line width = 0.1mm,->] (6,2) -- (7.4,2);
						%\draw[line width = 0.1mm,->] (6,2) -- (6,0.9);
						%\node at (6.7,1.8) {$e_1$};
						%\node at (5.8,1.4) {$e_2$};

						\node at (6.2,0) {$h$};
						\node at (2.7,-0.45) {$\ell = 1$};

						\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
						\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
						\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
						\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
						\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
						\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
						\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
						\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
						\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

						\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
						\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
						\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
						\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
						\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
						\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
						\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
						\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

						%\node at (3.2,-0.3) {$\ell = 1$};

					\end{tikzpicture}
				\end{center}
				\subcaption{Two-Dimensional Cantilever Beam}
			\end{minipage}
		}
	}
\end{figure}

In Chapter 1, the following parameters were defined. The parameter $h$ is the height of the two-dimensional beam and for the Timoshenko dimensionless constant $\displaystyle \alpha = \frac{A \ell^2}{I}$ was found.\\

For the Timoshenko beam model with a rectangular cross-section, the area moment of inertial is calculated as
\begin{eqnarray*}
	I = \frac{h^3b}{12}.
\end{eqnarray*}

After substitution, the following relationship between $\alpha$ and $h$ is found
\begin{eqnarray*}
	h = \sqrt{\frac{12}{\alpha}}.
\end{eqnarray*}
Using this relationship, $\alpha$ can be chosen in such a way that the beams have equal lateral dimensions.

\subsection{Accuracy of the numerical eigenvalues}
To calculate the eigenvalues and eigenfunctions of the Timoshenko beam, the authors of \cite{LVV09} use the article \cite{VV06}. This article was discussed in section \ref{sec:Timo:EigenvalueProblem} in this dissertation. Using this article, the eigenvalues for the Timoshenko beam model can be numerically calculated to a desired accuracy level.

A different approach is needed for the two-dimensional beam. The Finite Element Method is applied to the eigenvalue problem described in section \ref{sec:FEM:2D}. The following shows the rate of convergence of the first 20 eigenvalues using FEM and bi-cubic basis functions.

\begin{figure}[H]
	\begin{center}
		\scalebox{1.4}{
			\includegraphics[scale=0.7]{Convergence.png}
			\caption{Rate of convergence of the first 20 eigenvalues.}
			\label{fig:conv}
		}
	\end{center}	
\end{figure}
\textcolor{red}{FIX THIS}
Figure \ref{fig:conv} show the maximum relative error of the first 10 eigenvalues of the two-dimensional beam. In \cite{LVV09}, the authors use a grid size of $40 \times 10$ with $h= 1/10$ (or $\alpha = 1200$) and $h = 1/5$ (or $\alpha = 300$).\\

 Figure \ref{fig:conv} show that this can be improved significantly. For each case of $h$, the grid size is calculated as $n \times \max\left\{\lceil\frac{n}{h}\rceil,10\right\}$. This ensures that the nodes are distributed somewhat evenly relative to the length-width ratio. For this chapter, the results of the Timoshenko beam and two-dimensional beam are given accurate up to 5 significant digits.

\subsection{Comparing the shape of the eigenfunctions}
Before the eigenvalues can be compared, a method is required to be able to match the eigenvalues of the two different models. The authors of \cite{LVV09} match the modal shapes of the eigenfunctions. The corresponding eigenvalues can then be matched.\\

The eigenvalues of the Timoshenko beam, are referred to by the authors as beam-type eigenvalues.

\subsubsection{Shapes relating to beam-type eigenvalues}
The following figures are examples of beam-type mode shapes for the displacement $w$.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam1.jpg}
				\subcaption{2D Beam Type - $\lambda_6 = 21.911$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam2.png}
				\subcaption{Timoshenko - $\lambda_5 = 21.794$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam2.jpg}
				\subcaption{2D Beam Type - $\lambda_7 = 45.711$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam1.png}
				\subcaption{Timoshenko - $\lambda_6 = 45.390$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the beam-type 2D body and the Timoshenko beam with $\alpha = 4800$ ($h = 1/20$).}
		}
	}
\end{figure}
\FloatBarrier

\subsubsection{Shapes relating to non-beam type eigenvalues}
The following figures are examples of non beam-type mode shapes for the displacement $w$. These mode shapes are not present in the cantilever Timoshenko beam model and are not beam related. 
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam1.jpg}
				\subcaption{Non-Beam Type - $\lambda_4 = 7.7077$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam2.jpg}
				\subcaption{Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.}
		}
	}
\end{figure}
\FloatBarrier

The existence of these non-beam type eigenvalues imply that there are eigenvalues in the two-dimensional model, that are not in the Timoshenko beam model. This is an example of the complexities that arise when comparing models of different dimensions.

\subsubsection{Shapes of the cross-section}
The Timoshenko beam theory improve some one-dimensional beam theories such as the Euler-Bernoulli beam theory by also including the effect of shear. In the Timoshenko beam theory it is assumed that the cross-sections need not remain perpendicular to the neutral axis of the beam. The cross-sections however remain a straight line.\\

For the two-dimensional beam, the shape of the cross-section can deform into a s-shape. This is explained in \cite{LVV09} and a similar figure in \cite{LVV09} is given here.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=1.2]

	\draw[line width = 0.4mm] (0,1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (7,-1);
	\draw[line width = 0.4mm] (7,-1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (0,1);

	\draw[line width = 0.1mm,dashed] (3.5,1) -- (3.5,-1);

	\draw[thick] plot [smooth,tension=0.9] coordinates{(3.3,1) (3.7,0.5) (3.5,0) (3.3,-0.5) (3.7,-1)};

	\draw[line width = 0.2mm,red] (3.3,1) -- (3.7,-1);

\end{tikzpicture}
	\caption{S-shape deformation of the cross-section.}\label{fig:s}
\end{figure}

Note that figure \ref{fig:s} shows an exaggerated example of the deformation of a cross-section. The red line shows how the average rotation of a cross-section is calculated in \cite{LVV09}. This average rotation can be used to compare the rotation of the cross-section of the two-dimensional beam to the rotation of the cross-section of the Timoshenko beam.
\FloatBarrier

\subsubsection{Direct comparison of mode shapes}
Figure \ref{fig:w} shows the displacement of the material lines of the cantilever Timoshenko beam and the cantilever two-dimensional beam, compared directly.

\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2D.png}
			\caption{Comparison of the displacement $w$ mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:w}}}
\end{figure}
\FloatBarrier

Similarly Figure \ref{fig:phi} shows the direct comparison of the angle of the cross-section of the cantilever Timoshenko beam and the two-dimensional beam. The average rotation of the cross-section of the two-dimensional beam's mode shape is calculated as shown in figure \ref{fig:s} and \cite{LVV09}.

\FloatBarrier
\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2Dphi.png}
			\caption{Comparison of the angle $\phi$ (best fit for 2D Beam) mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:phi}}}
\end{figure}
\FloatBarrier

In \cite{LVV09} the authors obtained similar results for $\alpha=300$, which represents a short beam. Both figures \ref{fig:w} and \ref{fig:phi} were calculated with5 $\alpha = 4800$, which represents a longer, narrower beam.

\subsubsection{Remark}
The mode shapes were scaled to obtain figures \ref{fig:w} and \ref{fig:phi}. Therefore no scale is shown on the y-axis.

\subsection{Comparing the eigenvalues}
\subsubsection{Results of [LVV09]}
The following table contains the results form \cite{LVV09} as well as the corresponding results obtained in this dissertation.
\begin{table}[!ht]
	\centering
	\caption{Results from [LVV09] and results obtained in this dissertation. $0^a$ indicates a 0 as a result of rounding. $\alpha = 1200$}
	\begin{tabular}{|c|c|c|c||c|c|c|c|}
		\hline
		\multicolumn{4}{|c||}{Results from [LVV09]} & \multicolumn{4}{c|}{Dissertation} \\ \hline \hline
		~ & 2D & Timo & RE & ~ & 2D & Timo & RE  \\ \hline
		1 & 0.0317 & 0.0316 & 0.00315 & 1 & 0.031713 & 0.031639 & 0.0023407  \\ 
		2 & 1.14 & 1.14 & $0^a$ & 2 & 1.1413 & 1.1365 &  0.0042050 \\ 
		3 & 7.72 & - & ~ & 3 & 7.7161 & - &   \\ 
		4 & 7.92 & 7.86 & 0.00758 & 4 & 7.918 & 7.8617 & 0.0071116    \\ 
		5 & 26.2 & 25.9 & 0.0115 & 5 & 26.148 & 25.869 & 0.010669  \\ 
		6 & 60.8 & 59.9 & 0.0148 & 6 & 60.816 & 59.946 & 0.014303  \\ 
		7 & 69.3 & - & ~ & 7 & 69.344 & - &    \\ 
		8 & 115 & 113 & 0.0174 & 8 & 115.28 & 113.23 &  0.017787  \\ 
		9 & 192 & 188 & 0.0208 & 9 & 191.57 & 187.55 &  0.020999  \\ 
		10 & 192 & - &  & 10 & 192.03 & - &   \\ 
		11 & 291 & 284 & 0.0241 & 11 & 290.76 & 283.81 &  0.023899  \\ \hline
	\end{tabular}\label{Results_LVV09}
\end{table}

In Table \ref{Results_LVV09}, the results from \cite{LVV09} and the respective results from this dissertation is given. This table shows that the results of \cite{LVV09} can be replicated with more significant digits.\\

Interestingly, with the improved accuracy over \cite{LVV09}, the relative errors remain similar in Table \ref{Results_LVV09}. It is possible that the higher level of accuracy will only benefit larger eigenvalues. Larger eigenvalues are however not of interest for practical applications. The 2nd relative error of \cite{LVV09} however could not be duplicated.

\subsection{Comparison of the eigenvalues}
In \cite{LVV09}, two values for $\alpha$ are considered. For $\alpha = 1200$, the Timoshenko beam has a length to width ratio of $1:10$. For $\alpha = 300$, the Timoshenko beam has length to width ratio of $1:5$. These two values represent a ``long slender'' beam and a ``short'' beam.\\

In Table \ref{tab:Timo}, the eigenvalues of the Timoshenko beam are compared to the first 20 eigenvalues of the two-dimensional beam. The eigenvalues are calculated for different values of $\alpha$.\\

\begin{table}[h!]
	\scalebox{.8}{
	\makebox[\textwidth]{
	\caption{Eigenvalues of a Timoshenko cantilever beam vs the eigenvalues of a cantilever two-dimensional elastic body. RE - Relative Error}
	\begin{tabular}{|cccc||cccc||cccc||cccc|}
		\hline
		\multicolumn{16}{|c|}{Comparison of Eigenvalues} \\
		\hline\hline
		\multicolumn{4}{|c||}{ $h = 1/5$ or $\alpha = 300$}       & \multicolumn{4}{c||}{$h =1/10$ or $\alpha = 1200$}      & \multicolumn{4}{c||}{$h = 1/20$ or $\alpha = 4800$}      & \multicolumn{4}{c|}{$h = 1/30$ or $\alpha = 10800$} \\
		\hline
		{i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} \\
		\hline
		{1}&0.12151&1&0.12092&1&0.031713&1&0.031639&1&0.008013&1&0.008004&1&0.003568&1&0.003565\\
		{2}&3.5460&2&3.5071&2&1.1413&2     & 1.1365 & 2     & 0.30756 & 2     & 0.30705 & 2     & 0.13869 & 2     & 0.13855 \\
		\cellcolor{lightgray}{3} & \cellcolor{lightgray}{7.7311} &       & {-} & \cellcolor{lightgray}{3}     & \cellcolor{lightgray}{7.7161} &       & {-} & 3     & 2.3273 & 3     & 2.3213 & 3     & 1.0698 & 3     & 1.0683 \\
		{4} & 20.225 & 3     & 19.869 & 4     & 7.9180 & 3     & 7.8617 & \cellcolor{lightgray}{4}     & \cellcolor{lightgray}{7.7077} &       & {-} & 4     & 4.0140 & 4     & 4.0058 \\
		{5} & 56.109 & 4     & 54.766 & 5     & 26.148 & 4     & 25.869 & 5     & 8.5086 & 4     & 8.4762 & \cellcolor{lightgray}{5}     & \cellcolor{lightgray}{7.7047} &       & {-} \\
		\cellcolor{lightgray}{6} & \cellcolor{lightgray}{69.164} &       & {-} & 6     & 60.816 & 5     & 59.946 & 6     & 21.911 & 5     & 21.794 & 6     & 10.655 & 5     & 10.625 \\
		{7} & 114.03 & 5     & 110.75 & \cellcolor{lightgray}{7}     & \cellcolor{lightgray}{69.344} &       & {-} & 7     & 45.711 & 6     & 45.390 & 7     & 22.975 & 6     & 22.890 \\
		{8} & {189.17} &  6     & {186.50} & 8     & 115.28 & 6     & 113.23 & \cellcolor{lightgray}{8}     & \cellcolor{lightgray}{69.344} &       & {-} & 8     & 43.113 & 7     & 42.909 \\
		\cellcolor{lightgray}{9} & \cellcolor{lightgray}{192.61} &      &  & {9}     & {191.57} &   7  & {187.55} & 9     & 82.887 & 7     & 82.154 & \cellcolor{lightgray}{9}     & \cellcolor{lightgray}{69.331} &       & {-} \\
		{10} & 285.85 & 7     & 277.64 & \cellcolor{lightgray}{10}    & \cellcolor{lightgray}{192.03} &      & {-} & 10    & 136.03 & 8     & 134.58 & 10    & 73.230 & 8     & 72.803 \\
		{11} & 328.40 & 8     & 330.29 & 11    & 290.76 & 8     & 283.81 & \cellcolor{lightgray}{11}    & \cellcolor{lightgray}{192.48} &       & {-} & 11    & 115.41 & 9     & 114.61 \\
		\cellcolor{lightgray}{12} & \cellcolor{lightgray}{357.08} &       & {-} & \cellcolor{lightgray}{12}    & \cellcolor{lightgray}{374.45} &       & {-} & 12    & 207.29 & 9     & 204.69 & 12    & 171.61 & 10    & 170.20 \\
		{13} & 397.33 & 9     & 394.02 & 13    & 413.20 & 9     & 402.27 & 13    & 298.38 & 10    & 294.10 & \cellcolor{lightgray}{13}    & \cellcolor{lightgray}{192.52} &       & {-} \\
		{14} & 442.00   & 10    & 439.52 & 14    & 558.67 & 10    & 542.65 & \cellcolor{lightgray}{14}    & \cellcolor{lightgray}{376.83} &       & {-} & 14    & 243.56 & 11    & 241.26 \\
		\cellcolor{lightgray}{15} & \cellcolor{lightgray}{533.71} &       & {-} & \cellcolor{lightgray}{15}    & \cellcolor{lightgray}{614.11} &       & {-} & 15    & 410.63 & 11    & 404.01 & 15    & 332.83 & 12    & 329.28 \\
		{16} & 538.97 & 11    & 541.55 & 16    & 726.26 & 11    & 704.15 & 16    & 545.03 & 12    & 535.32 & \cellcolor{lightgray}{16}    & \cellcolor{lightgray}{377.16} &       & {-} \\
		\cellcolor{lightgray}{17} & \cellcolor{lightgray}{596.06} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{906.28} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{621.95} &       & {-} & 17    & 440.77 & 13    & 435.51 \\
		{18} & 602.77 & 12    & 596.09 & 18    & 913.69 & 12    & 884.92 & 18    & 702.30 & 13    & 688.64 & 18    & 568.51 & 14    & 561.04 \\
		\cellcolor{lightgray}{19} & \cellcolor{lightgray}{657.87} &       & {-} & 19    & 1113.7 & 13    & 1080.1 & 19    & 882.95 & 14    & 864.40 & \cellcolor{lightgray}{19}    & \cellcolor{lightgray}{623.05} &       & {-} \\
		{20} & 717.37 & 13    & 731.74 & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{1218.0}  &       & {-} & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{927.18} &       & {-} & 20    & 717.04 & 15    & 706.74 \\
		\hline
		\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{3.1718\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{3.1486\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.1018\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{1.4361\%} \\
		\hline
	\end{tabular}%
	\label{tab:Timo}%
}
}
\end{table}%
\FloatBarrier

The authors of \cite{LVV09} show that for $\alpha = 300$ the eigenvalues compare very well up to the 20th two-dimensional eigenvalue. This is also apparent for other values of $\alpha$ as Table \ref{tab:Timo} shows. In fact, the accuracy of the eigenvalues increases as $\alpha$ increases. The authors of \cite{LVV09} also mention that for $\alpha = 100$, the results are very good for the first 10 eigenvalues.\\

Interestingly, the number of non-beam type eigenvalues within the first few eigenvalues decreases as $\alpha$ increases (or more relevant for the two-dimensional beam: as $h$ decreases).\\

Overall, the relevant conclusions of \cite{LVV09} is verified. For practical applications of beam models, the Timoshenko beam can be used. Preferably for long slender beams as Table \ref{tab:Timo} shows. The authors also mention that nature of the disturbance should also be taken into consideration, mainly because the models are linear.\\

 The short comings of \cite{LVV09} is that damping is not considered. This is mentioned by the authors of \cite{LVV09}, but will not be addressed in this dissertation. Further work is to extend \cite{LVV09} to a three-dimensional comparison. This is investigated further in this chapter.\\

\section{Validity of the two-dimensional elastic model}
In this section, the validity of a two-dimensional cantilever beam is investigated. This is an extension of the work in \cite{LVV09}, as mentioned in the article.\\

As explained before, the two-dimensional model is used as a intermediate step between the Timoshenko and three-dimensional beam. The same method as in \cite{LVV09} will be used to investigate the validity of the two-dimensional model.\\

The models for the two-dimensional and three-dimensional beams are defined in section \ref{ssec:2D_Model:ModelProblem} and \ref{ssec:3D_Model:ModelProblems} respectively. These models are elastic bodies and not specifically beam models. Therefore it is expected to find eigenvalue and eigenfunction pairs that does not relate to beam type problems. This was already seen with the two-dimensional model in the previous section. These not beam type eigenvalues are calculated in this section, however the focus still remains on beam type problems.\\

\subsection{The models}
Recall a cantilever two-dimensional beam, referred to as Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem} and a cantilever three-dimensional beam, referred to as Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.\\

@@

For both models the parameter $h$ represents the width of the beams. The three-dimensional model also have depth $b$. Intuitively, from the figure, the two-dimensional beam should compare better to the three-dimensional beam for smaller $b$.

\subsection{Accuracy of the numerical eigenvalues}
Similar to how the eigenvalues and eigenfunctions of the two-dimensional model calculated in the previous section, the eigenvalues and eigenfunctions can be calculated for the three-dimensional model. The numerical formula to calculate the eigenvalues and eigenfunctions are derived in section \ref{sec:FEM:3D}.\\

Using piecewise tri-cubic basis functions, the following results were obtained.\\

\begin{tikzpicture}
	\begin{axis}[width=0.9\textwidth,
				 xlabel=Iteration Number,
				 ylabel=Error,
				 ymode=log,
				 legend pos=south east]
	\addplot[color=blue, mark=o] coordinates {
		(1,0.1)
		(2,0.05)
		(3,0.01)
		(4,0.005)
	};
	\addlegendentry{$h=1/5$}

	\addplot[color=red, mark=o] coordinates {
		(1,0.2)
		(2,0.1)
		(3,0.05)
		(4,0.02)
	};
	\addlegendentry{$h=1/10$}

	\addplot[color=green, mark=o] coordinates {
		(1,0.2)
		(2,0.1)
		(3,0.05)
		(4,0.02)
	};
	\addlegendentry{$h=1/20$}

	\addplot[color=black, mark=o] coordinates {
		(1,0.3)
		(2,0.3)
		(3,0.03)
		(4,0.03)
	};
	\addlegendentry{$h=1/30$}

	\end{axis}
	\end{tikzpicture}

\subsection{Comparing the mode shapes}
Identical to the method used in \cite{LVV09} and the previous section, the eigenvalues of the two models can be matched by comparing the mode shapes. As mentioned earlier, it is expected that there will be mode shapes not related to beam type problems, and also mode shapes that the three-dimensional model do not share with the two-dimensional model.\\

In this section, examples of the different mode shapes are shown.

@@

Figure \ref{fig:minipage11} looks similar to the beam type eigenvalues, however the shape is in the $x-z$ axis.\\

\subsection{Comparing the eigenvalues}
The eigenvalues can now be compared. In the tables of this subsection, the eigenvalues of the two-dimensional beam are compared to the eigenvalues of the three-dimensional beam. Since there are two parameters to consider, $h$ and $b$, only a few cases are considered.\\

The three-dimensional beam has more eigenvalues than the two-dimensional beam for the range considered in this investigation. Since the interest is in the comparison of the two models, the eigenvalues of the three-dimensional that are not present in the two-dimensional beam are omitted from the tables.\\

The eigenvalues for beam type problems, the main focus of this investigation, are highlighted in the tables.\\

The cases considered are similar to the cases in \cite{LVV09}, a short beam and a slender beam. For the short beam, $h = 1/5$ and for the slender beam, $h = 1/20$. The parameter $b$ will be varied to test the accuracy. For now the results are restricted to $b \leq h$. The other cases will be discussed later.\\

\subsubsection*{Case $b < h$:}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of a Two-dimensional beam} \label{sec:validity-of-a-cantilever-timoshenko-beam}
This section is an extention of section \ref{sec:validity-of-a-cantilever-timoshenko-beam} and the article \cite{LVV09}. As seen in the previous section, the two-dimensional model does have some characteristics that are not related to beam-type problems. This is also true for the three-dimensional model.

The results in this section will show that the three-dimensional model has a lot of non-beam type eigenvalues. It is due to this reason, the authors of \cite{LVV09} suggested to use the two-dimensional model as an intermediate step. The the previous section, the validity of a cantilever Timoshenko beam was investigated using a cantilever two-dimensional beam as a reference. In this section, the validity of a cantilever two-dimensional beam will be investigated using a cantilever three-dimensional beam as a reference.

\subsection{The models}
The two-dimensional model is the same model as used in the previous section, Problem T-2 defined in section \ref{ssec:2D_Model:ModelProblem}. From section \ref{ssec:3D_Model:ModelProblems}, the cantilever three-dimensional model, referred to as Problem 3D-1.

@@

\subsection{Calculating the eigenvalues}
Similar to calculating the eiegnvalues of the two-dimensional beam, to calculate the eigenvalues of the three-dimensional beam, the Finite Element Method is used. In section \ref{sec:FEM:3D}, the Finite Element Method for the three-dimensional beam is derived. Section \ref{3dFEM_EP} gives the following eigenvalue problem for the three-dimensional beam:

\subsubsection{Problem 3D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	K\bar{u} & = & M\lambda{\bar{u}},
\end{eqnarray} where $K$ and $M$ are the standard Finite Element Matrices for tri-cubic basis functions. The definition of the matrices are given in section \ref{sec:FEM:3D}.

\subsubsection{Accuracy of the eigenvalues}
Figure \ref{fig:conv_3d_eig} show the rate of convergence of the first 20 eigenvalues of Problem 3D-1E.

\begin{figure}[H]
    \centering
    \begin{adjustbox}{center}
        \includegraphics[scale=0.7]{Convergence.png}
    \end{adjustbox}
    \caption{Rate of convergence of the first 20 eigenvalues.}
    \label{fig:conv_3d_eig}
\end{figure}
\textcolor{red}{Update this figure.}

Similar to the two-dimensional case, the number of elements can be chosen so that at least the first 10 eigenvalues are accurate to 5 significant digits.

\subsection{Comparing the mode shapes}
The method to match up the eigenvalues remains the same as in the previous section. Since the three-dimensional model is more complex, there are more mode shapes that do no appear in the Timoshenko model or two-dimensional model. The focus remains on beam-type problems, so the mode shapes relating to beam type problems are most important.

@@

\subsection{Comparing the eigenvalues}
The eigenvalues of the two models can now be compared. Recall from the introduction, that the two-dimensional and three-dimensional models have different paramaters.

The two-dimensional model has the single paramater $h$ representing the height of the beam. The three-dimensional model has two parameters, $h$ and $b$, representing the height and width of the beam respectively.

\end{document}
